\documentclass{article}
\usepackage{booktabs, graphicx, hyperref, fontspec}
\usepackage{sectsty}
\allsectionsfont{\sffamily}
\usepackage[margin=1in]{geometry}
\hypersetup{
  colorlinks = true,
  urlcolor = cyan,
 }
 \providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
\usepackage[]{Fira Sans}

\begin{document}

\sffamily

\centerline{\Huge Game Theory}

\vspace{3 mm}

\centerline{\large Dr.~Ryan Safner}
\vspace{2 mm}
\centerline{\large \href{http://gameF21.classes.ryansafner.com}{gameF21.classes.ryansafner.com}}

\vspace{5 mm}

\begin{tabular}{@{}p{3.5in}p{3.5in}}           
\textbf{Course}: ECON 316 Fall
2021  & \textbf{Email:}  \href{mailto:safner@hood.edu}{\nolinkurl{safner@hood.edu}}\\
\textbf{Room}: Rosenstock 218 & \textbf{Office:}  Rosenstock 110\\
\textbf{Meets}: MW 11:30 AM---12:55PM & \textbf{Hours:}  MW 1:30
PM---2:30 PM \& by appt\\ 
\end{tabular}

\vspace{5 mm}

\hrule


\textbf{Game Theory} is a collection of analytical tools used to
rigorously study strategic interactions between rational agents. It is
one of the most useful and powerful theoretical innovations in
20\textsuperscript{th} century social science, and has been applied
extensively in economics, political science, psychology, sociology,
logic, computer science, animal behavior, and evolutionary biology. At
least twelve economists have won the Nobel Prize for research directly
related to game theory (and arguably more have made tangential
contributions). This is an \emph{economics} course, so the majority of
our focus and applications will be in economic settings, though we will
examine a much wider scope of strategic interactions than the ``narrow''
economic domain of firms and consumers.

Please don't get the wrong idea --- this course is \emph{not} the theory
of ``games'' in the colloquial sense (e.g.~chess, poker, football,
\emph{Fortnite}, etc)! This course is \emph{not} designed to make a you
a better {[}insert favorite game{]} player (but that may be a side
effect!). A ``game'' in social science is defined according to certain
technical criteria. Furthermore, formal game theory is \emph{highly
abstract and very mathematical}. I will try to balance formal theory
with interesting real world applications, sometimes including colloquial
``games.''

Game theory is a direct extension of basic microeconomics, so this class
assumes you have met the \textbf{required prerequisite course(s) ---
ECON 206}. My approach to game theory requires \emph{some} math, but it
will not be exclusively about the math.

\hypertarget{course-format-and-covid}{%
\section{Course Format and Covid}\label{course-format-and-covid}}

As of Fall 2021, all students are expected to be on campus except those
with special approved exemptions. As such, this course will be taught
\textbf{in-person} and \textbf{synchronously} until or unless otherwise
announced.

You are expected to come to class except due to medical reasons or other
legitimate conflicts. Watching videos are not a substitute for attending
class.\footnote{On average, even for students who complete all
  assignments, those that do not regularly attend class suffer by a full
  letter grade,
  (\href{https://www.aeaweb.org/articles?id=10.1257/jep.7.3.167}{Levitt
  1993}).} Please see the \protect\hyperlink{attendance}{attendance
policy} for more.

In any event that we are unable to meet in person, I will hold class
meetings at the same day/time live on Zoom, and post all recorded
lectures via Panopto on Blackboard, and all assignments will be
submitted online (often via Blackboard).

\hypertarget{learning-during-a-global-pandemic}{%
\subsection{Learning During a Global
Pandemic}\label{learning-during-a-global-pandemic}}

While we have made some progress in returning to normal, this remains a
unique semester and a lot of things are still awful right now. None of
us signed up for this. None of us are really okay, we're all just
pretending for everyone else.

Many of you may be dealing with hardships at home and at work, and are
generally juggling many more problems than usual. Everyone's future
plans have been completely put on hold or cancelled to a large degree.

I am prioritizing us supporting each other as human beings during this
crazy era, and will try to use simple, accessible solutions that make
sense for the most people, and above all, to be flexible.

If you tell me you're having trouble, I will do whatever I can to help,
and not judge you or think less of you. I hope you will extend me the
same courtesy.

You never \emph{owe} me personal information about your health (mental
or physical). You are however always welcome to talk to me about things
that you're going through. If I can't help you, I usually know somebody
who can.

I want you to learn a lot from this course, but it is more important for
you to remain healthy, balanced, and grounded during this crisis.

\hypertarget{course-objectives}{%
\section{Course objectives}\label{course-objectives}}

{By the end of this course,} you will:

\begin{itemize}
\tightlist
\item
  Recognize different types of strategic interactions across different
  domains (e.g.~economics, business, political science, biology, etc.)
\item
  Recognize common types of games, e.g.~prisoners' dilemma, stag hunt,
  battle of the sexes, chicken, hawk-dove
\item
  Solve for equilibria of games in normal form, extended form, in pure
  \& mixed strategies, with perfect and imperfect information
\item
  Understand the role of information, sequencing, credible commitments,
  repetition, etc.
\item
  Become familiar with some of the economics (and other) literatures
  that use game theoretic tools
\item
  Use game theory tools to explore a topic in writing
\end{itemize}

Given these objectives, this course fulfills all three of the learning
outcomes for
\href{https://www.hood.edu/academics/departments/george-b-delaplaine-jr-school-business/student-learning-outcomes}{the
George B. Delaplaine, Jr.~School of Business} Economics B.A. program:

\begin{itemize}
\tightlist
\item
  Use quantitative tools and techniques in the preparation,
  interpretation, analysis and presentation of data and information for
  problem solving and decision making {[}\ldots{]}
\item
  Apply economic reasoning and models to understand and analyze problems
  of public policy {[}\ldots{]}
\item
  Demonstrate effective oral and written communications skills for
  personal and professional success{[}\ldots{]}
\end{itemize}

{Fair warning:} \textbf{Economics is hard.} \emph{This, in particular,
may be of the hardest courses that you will take, primarily due to the
mathematical content.} Using the economic way of thinking is a skill, it
is literally retraining your brain to interpret and analyze the world in
a novel way, and is not something that can be memorized. I will do my
best to make this class intuitive and helpful, if not interesting. If at
any point you find yourself struggling in this course for any reason,
please come see me. Do not suffer in silence! Coming to see me for help
does not diminish my view of you, in fact I will hold you in
\emph{higher} regard for understanding your own needs and taking charge
of your own learning. There are also a some fantastic resources on
campus, such as the
\href{http://www.hood.edu/campus-services/academic-services/index.html}{Center
for Academic Achievement and Retention (CAAR)} and the
\href{http://www.hood.edu/library/}{Beneficial-Hodson Library}.

See my \href{/reference\#tips}{tips for success in this course}.

I reserve the right to change any part of this syllabus and course, at
my discretion, with proper advance warning.

\hypertarget{required-course-materials}{%
\section{Required Course materials}\label{required-course-materials}}

You can find all course materials at my \textbf{dedicated website} for
this course:
\href{https://gameF21.classes.ryansafner.com}{gameF21.classes.ryansafner.com}.
Links to the website are posted on our Blackboard course page. Please
familiarize yourself with the website, see that it contains this
\href{https://gameF21.classes.ryansafner.com/syllabus/}{syllabus},
guides for your
\href{https://gameF21.classes.ryansafner.com/reference/}{reference}, and
our \href{https://gameF21.classes.ryansafner.com/schedule/}{schedule}.
On the schedule page, you can find each module with its own class page
(\textbf{start there!}) along with all related readings, lecture slides,
practice problems, and assignments.

My lecture slides will be shared with you, and serve as your primary
resource, but our main ``textbook'' below is \textbf{recommended} as the
next best resource and will be available from the campus bookstore. I
will discuss more about textbooks and materials in the first module.

\hypertarget{books}{%
\subsection{Books}\label{books}}

My lecture slides (made available to you) are the primary resource for
the material and the best guide to prepare for homeworks and exams.
There are two books that we will roughly be following in parallel, both
available at the bookstore (or you can find them on Amazon, ebay, etc).
You may choose to purchase used or old versions, but be aware that
content and ordering may slightly vary across versions.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Dixit, Avinash K and Barry J Nalebuff, 2008, \emph{The Art of
  Strategy: A Game Theorist's Guide to Success in Business and Life},
  New York: W.W. Norton and Company
\item
  Dixit, Avinash K, Susan Skeath, and David H Reily Jr, 2014,
  \emph{Games of Strategy}, 4\textsuperscript{th} ed, New York: W.W.
  Norton and Company
\item
  Purchase a Moblab account (\$10) at
  \href{http://moblab.com}{moblab.com}
\end{enumerate}

The first book is a popular, read-on-the-beach style paperback that
covers the topics in a readable way. It is listed as \textbf{required}
in the bookstore, and should be fairly cheap on Amazon. It mirrors the
topics in the next book in roughly the same order but certainly at a
more general (and less mathematical) manner.

The second book is the more common textbook-style As with any textbook,
this will be more expensive but a deeper dive into the material that may
help with your understanding. It is listed as \textbf{recommended} in
the bookstore.

If you purchase the second book, it is superfluous for you to purchase
the first book, unless you really would like to!

Finally, purchase an account on \href{https://moblab.com}{moblab.com} so
that we can all play games in class together on common software.

I leave it to you to determine your own optimal strategy in purchasing
(or not) any textbooks. There is no dominant strategy here.

\hypertarget{assignments-and-grades}{%
\section{Assignments and Grades}\label{assignments-and-grades}}

Your final course grade is the weighted average of the following
assignments. You can find general descriptions for all the assignments
on the \href{assignments/}{assignments page} and more specific
information and examples on each assignment's page on the
\href{schedule/}{schedule page}.

\begin{tabular}{l|l|l}
\hline
Frequency & Assignment & Weight\\
\hline
n & Homeworks & 25\% (using average HW grade)\\
\hline
1 & Paper & 25\%\\
\hline
1 & Midterm Exam & 25\%\\
\hline
3 & Final Exam & 25\%\\
\hline
\end{tabular}

Each assignment is graded on a 100 point scale. Letter-grade equivalents
are based on the following scale:

\begin{table}
\centering
\begin{tabular}{l|c|l|c}
\hline
Grade & Range & Grade & Range\\
\hline
A & 93–100\% & C & 73–76\%\\
\hline
A− & 90–92\% & C− & 70–72\%\\
\hline
B+ & 87–89\% & D+ & 67–69\%\\
\hline
B & 83–86\% & D & 63–66\%\\
\hline
B− & 80–82\% & D− & 60–62\%\\
\hline
C+ & 77–79\% & F & < 60\%\\
\hline
\end{tabular}
\end{table}

See also my
\href{https://ryansafner.shinyapps.io/game_grade_calculator/}{
\texttt{Grade\ Calculator}} app where you can calculate your overall
grade using existing assignment grades and forecast ``what if''
scenarios.

These grades are firm cutoffs, but I do of course round upwards
\((\geq\) 0.5) for final grades. A necessary reminder, as an academic, I
am not in the business of \emph{giving} out grades, I merely report the
grade that you \emph{earn}. I will not alter your grade unless you
provide a reasonable argument that I am in error (which does happen from
time to time).

\textbf{No extra credit is available}

\hypertarget{policies-and-expectations}{%
\section{Policies and Expectations}\label{policies-and-expectations}}

This syllabus is a contract between you, the student, and me, your
instructor. It has been carefully and deliberately thought
out\footnote{A syllabus can and will be used as a legal document for
  disputes tried at a court of law. Ask me how I know.}, and I will
uphold my end of the agreement and expect you to uphold yours.

In the language of game theory, this syllabus is my commitment device. I
am a very understanding person, and I know that exceptions to rules
often need to be made for students. However, to be \emph{fair} to
\emph{all} students the syllabus artificially constrains my ability to
make exceptions at a whim for anyone. This prevents clever students from
exploiting my congenial personality at everyone else's expense. Please
read and familiarize yourself with the course policies and expectations
of you. Chances are, if you have a question, it is answered herein.

\hypertarget{attendence}{%
\subsection{Attendence}\label{attendence}}

Your day-to-day classroom attendance is not graded. My philosophy is
that you are all adults and must take ownership of your own learning or
else you will not succeed. Some assignments may require in-class
participation for credit, and an (unexcused) absence may be detrimental
to your grade. Attending class is one of the strongest predictors of
success.

However, as required under Hood College's ``Promise of Fall Plan,'' (Ch.
2-C) \textbf{your classroom attendance will be recorded at every class
meeting.} This is primarily to facilitate contact tracing.

If you know you will be absent, you are not \emph{required} to let me
know, but it is polite to give notice (Note if I do not reply to an
email of yours letting me know, I am probably busy but will still see it
and appreciate your email). Your absence will be noted and recorded for
the purposes stated above. If, however, we have an assignment due in
class, you \emph{must} notify me ahead of time in order to make
alternate arrangements to still receive credit. Hasty ex-post attempts
to notify me will generate little sympathy.

\hypertarget{late-assignments}{%
\subsection{Late Assignments}\label{late-assignments}}

I will accept late assignments, but will subtract a specified amount of
points as a penalty. Even if it is the last week of the semester, I
encourage you to turn in late work: some points are better than no
points!

\textbf{Homeworks}: If you turn in a homework after it is due but before
it is graded or the answer key posted, I generally will not take off any
points. However, \textbf{if you turn in a homework \emph{after} the
answer key is posted, I will automatically deduct 20 points (so the
maximum grade you can earn on it is an 80).}

\textbf{Exams}: If you know that you will be unable to complete an
\emph{in-class exam} as scheduled for a legitimate reason, please notify
me at least \emph{one week} in advance, and we will schedule a make-up
exam date. Failure to do so, including desperate attempts to make
arrangements only \emph{after} the exam will result in a grade of 0 and
little sympathy.

\textbf{Papers}: Starting at the deadline, I will take off 1 point for
every hour that your assignment is late.

I reserve the right to re-weight assignments for students whom I believe
are legitimately unable to complete a particular assignment.

\hypertarget{grading}{%
\subsection{Grading}\label{grading}}

I will try my best to post grades on Blackboard's Grading Center and
return graded assignments to you within about one week of you turning
them in. There will be exceptions. Where applicable, I will post answer
keys once I know most homeworks are turned in (see Late Assignments
above for penalties). Blackboard's Grading Center is the place to look
for your most up-to-date grades. See also my
\href{https://ryansafner.shinyapps.io/game_grade_calculator/}{
\texttt{Grade\ Calculator}} app where you can calculate your overall
grade using existing assignment grades and forecast ``what if''
scenarios.

\hypertarget{communication-email-slack-and-virtual-office-hours}{%
\subsection{Communication: Email, Slack, and Virtual Office
Hours}\label{communication-email-slack-and-virtual-office-hours}}

Students must regularly monitor their \textbf{Hood email accounts} to
receive important college information, including messages related to
this class. Email through the Blackboard system is my main method of
communicating announcements and deadlines regarding your assignments.
\textbf{Please do not reply to any automated Blackboard emails - I may
not receive it!}. My Hood email (\texttt{safner@hood.edu}) is the best
means of contacting me. I will do my best to respond within 24 hours. If
I do not reply within 48 hours, do not take it personally, and
\emph{feel free to send a follow up email} in the very likely event that
I genuinely did not see your original message.

Our \href{https://hoodcollegeeconomics.slack.com}{slack channel} is
available to all students and faculty in Economics and Business. I have
invited all of my classes and advisees. It will not be extended to
non-Business/Economics students or faculty. All users must use their
\textbf{hood emails} and \textbf{true first and last names}. Each course
has its own channel, exclusive for verified students in the course, and
myself, by my invite only. As a third party platform, you agree to its
Terms of Service. I have created this space as a way to stay connected,
to help one another, and to foster community. Behaviors such as posting
inappropriate content, harassing others, or engaging in academic
dishonesty, to be determined solely at my discretion, will result in one
warning, the content will be deleted, and subsequent behavior will
result in a ban.

In addition to in-person office hours, you can also make an appointment
for \textbf{``office hours''} on Zoom. You can join in with video,
audio, and/or chat, whichever you feel comfortable with. Of course, if
you are not available during those times, we can schedule our own time
if you prefer this method over email or Slack. If you want to go over
material from class, please have \emph{specific} questions you want help
with. I am not in the business of giving private lectures (particularly
if you missed class without a valid excuse).

Watch the excellent and accurate video explaining office hours
\href{https://gamef21.classes.ryansafner.com/syllabus/\#communication-email-slack-and-virtual-office-hours}{on
the website}.

\hypertarget{netiquette}{%
\subsection{Netiquette}\label{netiquette}}

When interacting with classmates and with your instructor online
(e.g.~Zoom, Slack, etc.) please follow appropriate internet etiquette
(``Netiquette''). Written communications lack important nonverbal cues
(such as body language, tone of voice, sarcasm, etc).

Above all else, please respect one another and think/reread carefully
about how others may see your post before you submit a comment. You are
expected to disagree and have different opinions, this is inherently
valuable in a discussion. Please be civil and constructive in responding
to others' comments: writing \emph{``have you considered `X'?''} is a
lot more helpful to all involved than just writing \emph{``well you're
just wrong.''}

Posting content that is willfully incendiary, illegal, or that
constitutes academic dishonesty (such as plagiarism) will automatically
earn a grade of 0 and may be elevated to other authorities on campus.

When using the chat function, please treat it as official course
communications, even though I may not be grading it. It may be a quick
and informal tool - don't feel you need to worry about spelling or
perfect grammar - but please try to avoid \emph{too} informal
``text-speak'' (i.e.~say ``That's good for you'' instead of ``thas good
4 u'').

\hypertarget{privacy}{%
\subsection{Privacy}\label{privacy}}

\href{https://www.execvision.io/blog/maryland-call-recording-laws/}{Maryland
law}
\href{https://law.justia.com/codes/maryland/2005/gcj/10-402.html}{requires}
all parties consent for a conversation or meeting to be recorded. If you
join in, and certainly if you participate, \textbf{you are consenting to
be recorded.} However, as described below, videos are \emph{not
accessible} beyond our class.

Live lectures are recorded on Zoom and posted to Blackboard via Panopto,
a secure course management system for video. Among other nice features
(such as multiple video screens, close captioning, and time-stamped
search functions!), Panopto is authenticated via your Blackboard
credentials, ensuring that \emph{our course videos are not accessible to
the open internet.}

For the privacy of your peers, and to foster an environment of trust and
academic freedom to explore ideas, \textbf{do not record our course
lectures or discussions.} You are already getting my official copies.

The
\href{https://www2.ed.gov/policy/gen/guid/fpco/ferpa/index.html}{Family
Educational Rights and Privacy Act} prevents me from disclosing or
discussing any student information, including grades and records about
student performance. If the student is at least 18 years of age,
\emph{parents (or spouses) do not have a right to obtain this
information}, except with consent by the student.

Many of you may be tuning in remotely, living with parents, and may have
occasional interruptions due to sharing a space. This is normal and
fine, but know that I will protect your privacy and not discuss your
performance when parents (or anyone other than you, for that matter) are
present, without your explicit consent.

\hypertarget{enrollment}{%
\subsection{Enrollment}\label{enrollment}}

Students are responsible for verifying their enrollment in this class.
The last day to add or drop this class with no penalty is
\textbf{Wednesday, September 1}. Be aware of
\href{https://www.hood.edu/offices-services/registrars-office/academic-calendar}{important
dates}.

\hypertarget{honor-code}{%
\subsection{Honor Code}\label{honor-code}}

Hood College has an Academic Honor Code which requires all members of
this community to maintain the highest standards of academic honesty and
integrity. Cheating, plagiarism, lying, and stealing are all prohibited.
All violations of the Honor Code are taken seriously, will be reported
to appropriate authority, and may result in severe penalties, including
expulsion from the college. See
\href{http://hood.smartcatalogiq.com/en/2016-2017/Catalog/The-Spirit-of-Hood/The-Academic-Honor-Code-and-Code-of-Conduct}{here}
for more detailed information.

\hypertarget{van-halen-and-mms}{%
\subsection{Van Halen and M\&Ms}\label{van-halen-and-mms}}

When you have completed reading the syllabus, email me a picture of the
band Van Halen and a picture of a bowl of M\&Ms.~If you do this
\emph{before} the date of the first exam, you will get bonus points on
the exam. If 75-100\% of the class does this, you each get 2 points. If
50-75\% of the class does this, you each get 4 points. If 25-50\% of the
class does this, you each get 6 points. If 0-25\% of the class does
this, you each get 8 points. Yes, you read this correctly.

\hypertarget{accessibility-equity-and-accommodations}{%
\subsection{Accessibility, Equity, and
Accommodations}\label{accessibility-equity-and-accommodations}}

College courses can, and should, be challenging and bring you out of
your comfort zone in a safe and equitable environment. If, however, you
feel at any point in the semester that certain assignments or aspects of
the course will be disproportionately uncomfortable or burdensome for
you due to any factor beyond your control, please come see me or email
me. I am a very understanding person and am happy to work out a solution
together. I reserve the right to modify and reweight assignments at my
sole discretion for students that I believe would legitimately be at a
disadvantage, through no fault of their own, to complete them as
described.

If you are unable to afford required textbooks or other resources for
any reason, come see me and we can find a solution that works for you.

This course is intended to be accessible for all students, including
those with mental, physical, or cognitive disabilities, illness,
injuries, impairments, or any other condition that tends to negatively
affect one's equal access to education. If at any point in the term, you
find yourself not able to fully access the space, content, and
experience of this course, you are welcome to contact me to discuss your
specific needs. I also encourage you to contact the
\href{https://www.hood.edu/academics/josephine-steiner-center-academic-achievement-retention/accessibility-services}{Office
of Accessibility Services} (301-696-3421). If you have a diagnosis or
history of accommodations in high school or previous postsecondary
institutions, Accessibility Services can help you document your needs
and create an accommodation plan. By making a plan through Accessibility
Services, you can ensure appropriate accommodations without disclosing
your condition or diagnosis to course instructors.

\hypertarget{tentative-schedule}{%
\section{Tentative Schedule}\label{tentative-schedule}}

\textbf{You can find a full schedule} with much more details, including
the readings, appendices, and other further resources for each class
meeting on the \href{schedule/}{schedule page}.

\end{document}