---
title: "Oligopoly Practice"
author: "Answer Key"
date: "ECON 316 — Game Theory — Fall 2021"
output:
  html_document:
    df_print: paged
    toc: true 
    toc_depth: 3
    toc_float: true
    code_folding: show
    highlight: tango

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,
                      warning = F,
                      message = F,
                      fig.retina = 3)
library(tidyverse)
update_geom_defaults("label", list(family = "Fira Sans Condensed"))
update_geom_defaults("text", list(family = "Fira Sans Condensed"))
```

Suppose Firm 1 and Firm 2 have a constant $MC=AC=8$. The market (inverse) demand curve is given by:

$$\begin{align*}
P&=200-2Q\\
Q&=q_1+q_2\\
\end{align*}$$

## Question 1 

#### Find the Cournot-Nash equilibrium output and profit for each firm.

Break demand apart into both firms' output. 
$$\begin{align*}
P&=200-2Q\\
P&=200-2(q_{1}+q_{2})\\
P&=200-2q_{1}-2q_2\\
\end{align*}$$

Solving for Firm 1, recalling that MR is twice the slope of the inverse demand curve:

$$MR_{1}=200-4q_{1}-2q_2$$


Firm 1 maximizes profit at $q^*$ where $MR=MC$:
$$\begin{align*}
MR_{1}&=MC && \text{Profit-max condition} \\
200-4q_{1}-2q_2&=8 && \text{Plugging in}\\
192-4q_1-2q_2&=0 && \text{Subtracting 8 from both sides}\\
192-2q_2&=4q_1 && \text{Adding }4q_{1} \text{ to both sides}\\
48-0.5q_1&=q_1^* && \text{Dividing both sides by 4}\\
\end{align*}$$

- Since Firm 2 is identical, its $q^*$ is:

$$q_2^*=48-0.5q_{1}$$

Find Nash equilibrium algebraically by plugging in one firm's reaction curve into the other's

$$\begin{align*}
q_{1}&=48-0.5q_2 && \text{Firm 1's reaction function}\\
q_{1}&=48-0.5(48-0.5q_{1}) && \text{Plugging in Firm 2's reaction function}\\
q_{1}&=48-24+0.25q_{1} && \text{Distributing the }-0.5\\
q_{1}&=24+0.25q_{1} && \text{Subtracting}\\
0.75q_{1}&=24 && \text{Subtracting }0.25q_{1} \text{ to both sides}\\
q_{1}&=32 && \text{Dividing by 0.75}\\
\end{align*}$$

Symmetrically, $q_{2}=32$

Both countries produce 32:

$$\begin{align*}
P&=200-2(32)-2(32)\\
P&=\$72\\
\end{align*}$$

We can find the profit for each firm:

$$\begin{align*}
\pi_{1}&=q_{1}(P-c)\\
\pi_{1}&=32(72-8)\\
\pi_{1}&=\$2,048\\
\end{align*}$$

## Question 2 

#### Find the output and profit for each firm if the two were to collude.

Suppose now both firms **collude** and act like a single monopolist, who sets:

$$\begin{align*}
MR&=MC && \text{Profit-max condition}\\
200-4Q&=8 && \text{Plugging in}\\
192-4Q&=0 && \text{Subtracting 8 from both sides}\\
192&=4Q && \text{Adding }4Q \text{ to both sides}\\
48&=Q && \text{Dividing both sides by 4}\\
\end{align*}$$

So each firm will produce 24. 

The monopoly price will then be

$$\begin{align*}
P&=200-2Q\\
P&=200-2(48)\\
P&=\$104\\
\end{align*}$$

Total profit will then be:

$$\begin{align*}
\Pi&=Q(P-c)\\
\Pi&=48(104-8)\\
\Pi&=\$4,608\\
\end{align*}$$

with $2,304 going to each firm

## Question 3

#### Find the price and output if the two were to compete on price instead of quantity.

In Bertrand competition, each firm sets:

$$\begin{align*}
P&=MC\\
200-2Q&=8\\
192-2Q&=0\\
192&=2Q\\
96&=Q\\
\end{align*}$$

Each firm produces 48, earning no profits.

## Question 4

#### Suppose Firm 1 produces before Firm 2, who can see how much Firm 1 produced. Find the price, output, and profit for each firm.

Substitute follower's reaction function into market (inverse) demand function

$$\begin{align*}
P&=200-2q_{1}-2q_2 && \text{The inverse market demand function}\\
P&=200-2q_{1}-2(48-0.5q_{1}) && \text{Plugging in Firm 2's reaction function for} q_2\\
P&=200-2q_{1}-96+1q_{1} && \text{Multiplying by }-3\\
P&=104-q_{1} && \text{Simplifying the right}\\
\end{align*}$$

- Find $MR$ for Firm 1 from market demand

$$MR_{1}=104-2q_{1}$$

$$\begin{align*}
MR&=MC && \text{Profit-max condition}\\
104-2q_{1}&=8 && \text{Plugging in}\\
104&=8+2q_{sa} && \text{Adding }2q_{1} \text{ to both sides}\\
96&=2q_{sa} && \text{Subtracting 20 from both sides}\\
48&=q_{sa}^* && \text{Dividing both sides by 2} \\
\end{align*}$$

- Firm 2 will respond

$$\begin{align*}
q_2^*&=48-0.5q_{1}\\
q_2^*&=48-0.5(48)\\
q_2^*&=48-24\\
q_2^*&=24\\
\end{align*}$$

- With $q^*_{1}=48$ and $q^*_2=24$, this sets a market price of

$$\begin{align*}
P&=200-2Q\\
P&=200-2(72)\\
P&=56\\
\end{align*}$$

- Profit for Firm 1 is 

$$\begin{align*}
\pi_{1}&=q_{1}(P-c)\\
\pi_{1}&=48(56-8)\\
\pi_{1}&=\$2,304\\
\end{align*}$$

- Profit for Firm 2 is 

$$\begin{align*}
\pi_{2}&=q_{2}(P-c)\\
\pi_{2}&=24(56-8)\\
\pi_{2}&=\$1,152\\
\end{align*}$$

```{r}
library(ggtext)
f1=function(x){48-0.5*x}
f2=function(x){96-2*x}

points<-tribble(
  ~x, ~y, ~lab,
  32, 32, "Cournot Nash Eq.",
  24, 48, "Stackelberg Nash Eq. (Firm 1 leads)",
  24, 24, "Collusion"
)

ggplot(tibble(x=c(0,10)), aes(x=x))+
  stat_function(fun=f1, geom="line", size=2, color="red")+
    annotate(geom = "label", x = 45, y = f1(45) , label = "Firm 1's BR",color = "red", size = 4) +
  stat_function(fun=f2, geom="line", size=2, color="blue")+
    annotate(geom = "label", x = 45, y = f2(45) , label = "Firm 2's BR",color = "blue", size = 4) +
  geom_segment(x = 0, xend = 32, y = 32, yend = 32, size = 1, linetype = "dotted")+
  geom_segment(x = 32, xend = 32, y = 32, yend = 0, size = 1, linetype = "dotted")+
  
  geom_segment(x = 0, xend = 24, y = 48, yend = 48, size = 1, linetype = "dotted")+
  geom_segment(x = 24, xend = 24, y = 48, yend = 0, size = 1, linetype = "dotted")+
  
  geom_segment(x = 0, xend = 24, y = 24, yend = 24, size = 1, linetype = "dotted")+
  geom_segment(x = 24, xend = 24, y = 24, yend = 0, size = 1, linetype = "dotted")+
  geom_point(data = points,
             aes(x = x,
                 y = y),
             size = 3)+
  ggrepel::geom_label_repel(data = points,
                           aes(x = x,
                               y = y,
                               label = lab),
                           seed = 2,
                           box.padding = 0.5)+
  scale_x_continuous(breaks = seq(0,100,10),
                     limits = c(0,100),
                     expand=c(0,0))+
  scale_y_continuous(breaks = seq(0,100,10),
                     limits = c(0,100),
                     expand=c(0,0))+
  theme_classic(base_family = "Fira Sans Condensed", base_size=16)+
  labs(x = "<span style='color:#0047AB;'>Firm 2's output</span>",
       y = "<span style='color:#D7250E;'>Firm 1's output</span>",
       title = "Firms' Reaction Curves")+
  theme(axis.title.x = element_markdown(),
        axis.title.y = element_markdown())
```

```{r}
Demand_market<-function(x){200-2*x}
MR_market<-function(x){200-4*x}
MC_market<-function(x){8}

CS_1<-tribble(
  ~x, ~y,
  0, 200,
  0, 80,
  40, 80
)

DWL_1<-tribble(
  ~x, ~y,
  40, 20,
  40, 80,
  60, 20
)

points_m <- tribble(
  ~x, ~y, ~lab,
  64, 72, "Cournot Nash Eq.",
  48, 104, "Collusion",
  96, 8, "Bertrand/P.C.",
  72, 56, "Stackelberg Nash Eq."
)
ggplot(tibble(x=c(0,10)), aes(x=x))+
  #geom_polygon(data=CS_1, aes(x=x,y=y), fill="blue",alpha=0.5)+ #cs
  #geom_rect(xmin=0, xmax=40, ymin=20, ymax=80, fill="green", alpha=0.5)+ # pi
  #geom_polygon(data=DWL_1, aes(x=x,y=y), fill="black",alpha=0.5)+ #dwl
  stat_function(fun=Demand_market, geom="line", size=2, color="blue")+
  annotate(geom = "label", x = 30, y = Demand_market(30) , label = "Market Demand",color = "blue", size = 4) +

  stat_function(fun=MR_market, geom="line", size=2, color="purple")+
  annotate(geom = "label", x = 15, y = MR_market(15) , label = "MR(q)",color = "purple", size = 4) +

  stat_function(fun=MC_market, geom="line", size=2, color="red")+

  geom_segment(x=64,xend=64,y=0,yend=72,linetype="dotted", size =1)+
  geom_segment(x=0,xend=64,y=72,yend=72,linetype="dotted", size =1)+

  geom_segment(x=48,xend=48,y=0,yend=104,linetype="dotted", size =1)+
  geom_segment(x=0,xend=48,y=104,yend=104,linetype="dotted", size =1)+
  
  geom_segment(x=96,xend=96,y=0,yend=8,linetype="dotted", size =1)+
  #geom_segment(x=0,xend=96,y=8,yend=8,linetype="dotted", size =1)+

  geom_segment(x=72,xend=72,y=0,yend=56,linetype="dotted", size =1)+
  geom_segment(x=0,xend=72,y=56,yend=56,linetype="dotted", size =1)+

  annotate(geom = "label", x = 70, y = MC_market(70) , label = "MC(q)=AC(q)",color = "red", size = 4) +

  geom_point(data = points_m,
             aes(x = x,
                 y = y),
             size = 3)+
  ggrepel::geom_label_repel(data = points_m,
                           aes(x = x,
                               y = y,
                               label = lab),
                           seed = 2,
                           box.padding = 0.75)+

  scale_x_continuous(breaks = seq(0,100,10),
                     limits = c(0,105),
                     expand=c(0,0))+
  scale_y_continuous(breaks = seq(0,200,20),
                     labels = scales::dollar,
                     limits = c(0,200),
                     expand=c(0,0))+
  theme_classic(base_family = "Fira Sans Condensed", base_size=16)+
  labs(x = "Industry Output, Q",
       y = "Market Price, p",
       title = "Industry Equilibria")
```
